---
title: "Assignment"
output:
  pdf_document: default
  html_document: default
date: '2024-11-28'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r load_libraries, echo=FALSE}
library(readxl)      # For reading Excel files
library(forecast)
library(ggplot2)
library(keras)
library(tensorflow)
library(tseries)
library(dplyr)

```

```{r load_data, echo=FALSE}
elec_train <- read_excel('2023-11-Elec-train.xlsx') 
colnames(elec_train) <- c("Date", "Power_kw", "Temp_c")

elec_train$Date <- as.POSIXct(elec_train$Date, format="%m/%d/%Y %H:%M", tz="UTC")
# Add time features
elec_train$Hour <- as.numeric(format(elec_train$Date, "%H"))
elec_train$Day <- as.numeric(format(elec_train$Date, "%u"))


head(elec_train)


```


```{r check_missing_value, echo=FALSE}

colSums(is.na(elec_train))
```



### Handle missing values 

Plot Missing Data for power 
```{r missing_data_power, echo= FALSE}

library(imputeTS)

ggplot_na_distribution(elec_train$Power_kw)


elec_train$Power_kw = na_interpolation(elec_train$Power_kw)

ggplot_na_distribution(elec_train$Power_kw)

```
```{r check_missing_value_2, echo=FALSE}

colSums(is.na(elec_train))
```


```{r check_missing_value_3, echo=FALSE}

colSums(is.na(elec_train))
```
```{r missing_value_after_dateformat, echo=FALSE}
# Remove rows where Date is NA
elec_train <- elec_train[!is.na(elec_train$Date), ]

# Verify that no missing values remain in the Date column
colSums(is.na(elec_train))
```
## Summmry 

```{r summary, echo=FALSE}
summary(elec_train)
```

## Including Plots

You can also embed plots, for example:

```{r plot_data, echo=FALSE}
plot(elec_train)
```


### ACF , PACF and ADF Test

To check the Seasonality for Power- ACF AND PACF

```{r check_seasonality_power}

# Convert to time series
power_ts <- ts(elec_train$Power_kw, frequency = 96)

# ACF and PACF to check seasonality for power
acf(power_ts, lag.max = 100)
pacf(power_ts, lag.max = 100)

```


To check the Seasonality for Temperature - ACF AND PACF
```{r seasonality_check_decompose, echo=FALSE}
print(frequency(power_ts))

decomposition <- decompose(power_ts, type = "additive")
plot(decomposition)

```
We could see the stron seaonality in the data. 


Test for Stationarity

```{r test_stationrity_power, echo=FALSE}

library(tseries)

adf_test <- adf.test(elec_train$Temp_c, alternative = "stationary")
print(adf_test)

```

Above concludes that we can make ARIMA test


ACF: Significant autocorrelation at lag 1 and several subsequent lags, indicating potential for ARIMA models.
PACF: A sharp drop after lag 1, which suggests an AR(1) process could be appropriate.


### rTime Seies Plot

```{r time series plotr, echo=FALSE}
library(ggplot2)
library(forecast)

autoplot(power_ts) +
  ggtitle("Electricity Consumption Time Series") +
  xlab("Time") +
  ylab("Power (kW)")

```
```{r }
fit=tslm(power_ts~Temp_c+Hour+Day+trend+season,data=elec_train)
summary(fit)
```
Here we could see that Temperature plays a significant role in the power consumption 


```{r}
CV(fit)
```
```{r}
checkresiduals(fit, test=FALSE, plot=TRUE)
```
```{r}
checkresiduals(fit, test='LB', plot=FALSE)
```

```{r}
Box.test(residuals(fit), lag = 10, type = "Ljung-Box")

acf(residuals(fit))
```


```{r}
library(testcorr)

cc.test(elec_train[,"Power_kw"],elec_train[,"Temp_c"],max.lag = 1, plot = FALSE)
```
#Split the dataset for traububg

```{r split_train_test, echo=FALSE}
set.seed(666) 
train <- power_ts[1:(length(power_ts) - 96)]
test <- power_ts[(length(power_ts) - 95):length(power_ts)]

# Verify splits
cat("Training set length:", length(train), "\n")
cat("Testing set length:", length(test), "\n")


if (length(train) < 192) {
  stop("Training data must contain at least two seasonal cycles (192 observations for daily seasonality).")
}

```


##  Forecasting Without Temperature


### Holt-Winter Mode 

```{r holt_winter_without_temp_test, echo=FALSE}

# Ensure train is a ts object with correct frequency
train <- ts(train, frequency = 96)

# Check if training set has at least two full seasonal cycles
if (length(train) >= 192) {
  # Fit Holt-Winters with additive seasonality
  hw_add <- HoltWinters(train, seasonal = "additive")
  hw_add_forecast <- forecast(hw_add, h = length(test))
  
  # Calculate RMSE
  rmse_hw_add <- sqrt(mean((test - hw_add_forecast$mean)^2))
  cat("Holt-Winters Additive RMSE:", rmse_hw_add, "\n")
} else {
  stop("Training set must contain at least two full seasonal cycles (192 observations).")
}




```



### SARIMA 

``` {r SARIMA_without_temp_test, echo=FALSE}
library(forecast)

# Fit SARIMA
sarima_model <- auto.arima(train, seasonal = TRUE)

# Forecast for the test period
sarima_forecast <- forecast(sarima_model, h = length(test))

# Calculate RMSE
rmse_sarima <- sqrt(mean((test - sarima_forecast$mean)^2))
cat("SARIMA RMSE:", rmse_sarima, "\n")

```



## With Temperature


### SARIMA with tempertaure
``` {r Split_with_temperature, echo=FALSE}


# Ensure temperature data is aligned with the time series
train_with_temp <- elec_train[1:(nrow(elec_train) - 96), c("Power_kw", "Temp_c")]
test_with_temp <- elec_train[(nrow(elec_train) - 95):nrow(elec_train), c("Power_kw", "Temp_c")]


```

``` {r SARIMA_with_temp_test, echo=FALSE}


library(forecast)

# Fit ARIMA with temperature as an exogenous variable
sarima_with_temp <- auto.arima(train_with_temp$Power_kw, xreg = train_with_temp$Temp_c)

# Forecast using temperature for the test period
sarima_temp_forecast <- forecast(sarima_with_temp, xreg = test_with_temp$Temp_c, h = length(test_with_temp$Power_kw))

# Calculate RMSE
rmse_sarima_temp <- sqrt(mean((test_with_temp$Power_kw - sarima_temp_forecast$mean)^2))
cat("SARIMA RMSE (with temperature):", rmse_sarima_temp, "\n")



```

### Random Forest with tempertaure
``` {r randomforest_with_temp_test, echo=FALSE}

library(randomForest)

# Prepare lagged features and temperature for Random Forest
train_rf <- data.frame(
  Lag1 = lag(train_with_temp$Power_kw, 1),
  Temp = train_with_temp$Temp_c,
  Target = train_with_temp$Power_kw
) %>% na.omit()  

test_rf <- data.frame(
  Lag1 = lag(test_with_temp$Power_kw, 1),
  Temp = test_with_temp$Temp_c
) %>% na.omit()

# Train the Random Forest model
rf_model <- randomForest(Target ~ Lag1 + Temp, data = train_rf, ntree = 100)

# Forecast using Random Forest
rf_forecast <- predict(rf_model, test_rf)

# Calculate RMSE
rmse_rf <- sqrt(mean((test_with_temp$Power_kw[2:length(test_with_temp$Power_kw)] - rf_forecast)^2))
cat("Random Forest RMSE (with temperature):", rmse_rf, "\n")


```


## Result 


```{r print_results}
cat("Final Results:\n")
cat("Holt-Winters RMSE (without temperature):", rmse_hw_add, "\n")
cat("SARIMA RMSE (without temperature):", rmse_sarima, "\n")

cat("SARIMA RMSE (with temperature):", rmse_sarima_temp, "\n")
cat("Random Forest RMSE (with temperature):", rmse_rf, "\n")
```
### VAR 

```{r}
library(fpp2)
```

```{r}
library(vars)
set.seed(667)  # For 
var_model <- VAR(train_with_temp, p=1, type = "const", season = 96)

```

```{r}
summary(var_model)
```

```{r}
# Forecast for the next 96 days (96 * 24 * 4 = 9216 intervals)
forecast_results <- predict(var_model, n.ahead = 9216)
```

```{r}
# Extract the forecasted Power values
power_forecast <- forecast_results$fcst$Power_kw[, "fcst"]

```

```{r}
# Calculate RMSE
actual_power <- test_with_temp$Power_kw
var_rmse <- sqrt(mean((actual_power - power_forecast)^2))

# Calculate Mean of the Forecast
forecast_mean <- mean(power_forecast)

# Print RMSE and Mean
cat(" VAR RMSE (With Temperature):", var_rmse, "\n")
cat("Mean of the Forecasted Power:", forecast_mean, "\n")


# Save the forecast results to a CSV
write.csv(power_forecast, "VAR_power_forecast_96_days.csv", row.names = FALSE)

# Plot the forecast
plot.ts(power_forecast, main = "Forecasted Power for Next 96 Days (15-min intervals)", ylab = "Power (kW)", xlab = "Time")

```
### XG Boost


```{r xgboost, echo=FALSE}
# Install required packages
library(xgboost)

# Prepare data for XGBoost
train_size <- floor(0.8 * nrow(elec_train))
train_data <- elec_train[1:train_size, ]
test_data <- elec_train[(train_size + 1):nrow(elec_train), ]

# Prepare training and testing features (X) and target (y)
train_x <- as.matrix(train_data[, c("Temp_c", "Hour", "Day")])  # Features
train_y <- train_data$Power_kw  # Target variable

test_x <- as.matrix(test_data[, c("Temp_c", "Hour", "Day")])  # Features
test_y <- test_data$Power_kw  # Target variable

# Convert to DMatrix format
dtrain <- xgb.DMatrix(data = train_x, label = train_y)
dtest <- xgb.DMatrix(data = test_x, label = test_y)

# Train an XGBoost model
xgb_model <- xgb.train(
  data = dtrain, 
  nrounds = 100, 
  objective = "reg:squarederror"
)

# Predict using XGBoost
xgb_predictions <- predict(xgb_model, newdata = dtest)

# Evaluate XGBoost performance
rmse_xgb <- sqrt(mean((test_y - xgb_predictions)^2))
cat("XGBoost RMSE:", rmse_xgb, "\n")

```


```{r}
# Compute RMSE for VAR and SARIMA
rmse_var <- sqrt(mean((test_y - power_forecast)^2))


cat("Final Results:\n")
cat("Holt-Winters RMSE (without temperature):", rmse_hw_add, "\n")
cat("SARIMA RMSE (without temperature):", rmse_sarima, "\n")

cat("SARIMA RMSE (with temperature):", rmse_sarima_temp, "\n")
cat("Random Forest RMSE (with temperature):", rmse_rf, "\n")
cat("VAR RMSE (with temperature):", rmse_var, "\n")

cat("XGBoost RMSE ( with temperature):", rmse_xgb, "\n")

```


```{r}
library(openxlsx)
aligned_sarima_forecast <- sarima_forecast$mean[2:length(sarima_forecast$mean)]

predictions <- data.frame(
  Without_Temperature = aligned_sarima_forecast,  # SARIMA predictions
  With_Temperature = rf_forecast              # Random Forest predictions
)

# Export the predictions to an Excel file
output_file <- "janani_sinniah_vellaswamy.xlsx"
write.xlsx(predictions, output_file, row.names = FALSE)

cat("Predictions exported to:", output_file, "\n")
```
